### 2020-02-20 initial/notes.tex
fix sign error, actually plot tidal breakdown criterion

    \item $\omega_{s1}$ comes from the action of the central star on the
        quadrupole moment of the planet. This comes from applying the
        quadrupolar torque to the spin angular momentum of the planet, or
        \begin{align}
            \omega_{s1} &\sim \frac{GM_* J_2 m_1 R_1^2/a^3}{k_p m_1 R_1^2s},\\
                &\sim \frac{k_{qp}}{k_p}\frac{M_*}{m_1} \p*{\frac{R_1}{a}}^3.
        \end{align}
        Note that $J_2 = \frac{k_{qp}s^2}{Gm_1/R_1^3}$, some constant times the
        spin frequency relative to breakup. This captures the scalings.

@@

Note that for Jupiter, $k_p \sim 0.27$ while $J_2 \sim \scinot{1.4}{-2}$ giving
$k_{qp} \sim 0.0467$ since $k_{qp} \frac{\Omega}{\sqrt{GM_p/R_p^3}} = J_2$.
Consider again the precession frequencies
\begin{align}
    \omega_{s1} &= \frac{3k_{qp}}{2k_p}\frac{M_*}{m_1}
            \p*{\frac{R_1}{a_1}}^3s,\\
        &\approx 10\;\mathrm{yr^{-1}}
            \p*{\frac{M_*/m_1}{300}}
            \p*{\frac{5.5 k_{qp}}{k_p}}\p*{\frac{R_1}{2R_J}}^3
                \p*{\frac{a_1}{0.05\;\mathrm{AU}}}^{-3}
                \p*{\frac{s}{2\pi / 10\;\mathrm{hr}}},\\
    \omega_{1p} &= \frac{3m_p}{4m_1} \frac{a_1}{a_p\p*{1 - e_p^2}} \Omega_1,\\
        &\approx 0.0625\;\mathrm{yr^{-1}}
            \frac{m_p/m_1}{0.003}
            \p*{\frac{a_1/a_p}{0.05}}
            \frac{\Omega_1}{2\pi / 4\;\mathrm{days}}
            \frac{1}{1 - e_p^2}.
\end{align}
Let's note that the effective initial $\eta = \frac{\omega_{1p}\cos
I}{\omega_{s1}} \sim 0.006$. Furthermore, the smallest $\eta$ we can get at
small final obliquities is just when $s = \Omega_1$, which is roughly a factor
of $10$ increase so $\eta_f \sim 0.06$. Recall that $\eta = \frac{s_c}{s}
\approx \frac{s_c}{\Omega_1}0.1$ since $\Omega_1/s \approx 0.1$, then $s_c
\approx 0.06$ as well.

@@

Let's return to the simple EOM considered in the constant $\eta$ tides case,
where
\begin{align}
    \pd{\phi}{t} &= -\mu + \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
We want to determine where tides destroys CS2. We saw before that tides tends to
shift CS2/CS4 towards each other. In fact, we can immediately see where they
will collide: in the $\dot{\mu}$ equation, no $\phi$ is stationary if
\begin{equation}
    \epsilon > \frac{\eta \sin I}{1 - \mu^2}.
\end{equation}
Note that we can invert this to find, for a given $\epsilon$, the smallest value
of $\eta$ for which CS2 is still stable, which must satisfy $\eta \sin I =
\epsilon\p*{1 - \eta^2\cos^2 I}$, or
\begin{equation}
    \eta_{\min} \equiv \frac{\sin I + \sqrt{\sin^2I
        + 4\epsilon^2 \cos^2 I}}{2\epsilon \cos^2 I}.
\end{equation}

>>>>

    \item $\omega_{s1}$ comes from the action of the central star on the
        quadrupole moment of the planet. This comes from applying the
        quadrupolar torque to the spin angular momentum of the planet, or
        \begin{align}
            \omega_{s1} &\sim \frac{GM_* J_2 m_1 R_1^2/a^3}{k_p m_1 R_1^2s},\\
                &\sim \frac{k_{qp}}{k_p}\frac{M_*}{m_1} \p*{\frac{R_1}{a}}^3s.
        \end{align}
        Note that $J_2 = \frac{k_{qp}s^2}{Gm_1/R_1^3}$, some constant times the
        spin frequency relative to breakup. This captures the scalings.

@@

Note that for Jupiter, $k_p \sim 0.27$ while $J_2 \sim \scinot{1.4}{-2}$ giving
$k_{qp} \sim 0.0467$ since $k_{qp} \frac{\Omega}{\sqrt{GM_p/R_p^3}} = J_2$.
Consider again the precession frequencies (we will take a hot Jupiter at
$0.1\;\mathrm{AU}$ and a cold Earth at $1\;\mathrm{AU}$, roughly)
\begin{align}
    \omega_{s1} &= \frac{3k_{qp}}{2k_p}\frac{M_*}{m_1}
            \p*{\frac{R_1}{a_1}}^3s,\\
        &\approx 0.062\;\mathrm{yr^{-1}}
            \p*{\frac{M_*/m_1}{300}}
            \p*{\frac{5.5 k_{qp}}{k_p}}\p*{\frac{R_1}{R_J}}^3
                \p*{\frac{a_1}{0.1\;\mathrm{AU}}}^{-9/2}
                \p*{\frac{s/\Omega_1}{10}},\\
    \omega_{1p} &= \frac{3m_p}{4m_1} \p*{\frac{a_1}{a_p\p*{1 - e_p^2}}}^3
            \Omega_1,\\
        &\approx 0.0047\;\mathrm{yr^{-1}}
            \frac{m_p/m_1}{0.003}
            \p*{\frac{a_1/a_p}{0.1}}
            \frac{\Omega_1}{2\pi / 11.54\;\mathrm{days}}
            \frac{1}{\p*{1 - e_p^2}^3}.
\end{align}
Note then that $s_c$ is where $\omega_{s1} = \omega_{1p}$, which is satisfied
for $s_c \approx 0.76$.

@@

Let's return to the simple EOM considered in the constant $\eta$ tides case,
where
\begin{align}
    \pd{\phi}{t} &= -\mu + \eta\p*{\cos I + \sin I \frac{\mu}{\sqrt{1 - \mu^2}}
        \cos \phi} ,\\
    \pd{\mu}{t} &= -\eta \sin I \sin \phi + \s*{\epsilon \p*{1 - \mu^2}},
\end{align}
We want to determine where tides destroys CS2. We saw before that tides tends to
shift CS2/CS4 towards each other. In fact, we can immediately see where they
will collide: in the $\dot{\mu}$ equation, no $\phi$ is stationary if
\begin{equation}
    \epsilon > \frac{\eta \sin I}{1 - \mu^2}.
\end{equation}
Note that we can invert this to find, for a given $\epsilon$, the smallest value
of $\eta$ for which CS2 is still stable, which must satisfy $\eta \sin I =
\epsilon\p*{1 - \eta^2\cos^2 I}$, or
\begin{equation}
    \eta_{\min} \equiv \frac{-\sin I + \sqrt{\sin^2I
            + 4\epsilon^2 \cos^2 I}}{2\epsilon \cos^2 I}
        \approx \frac{2\epsilon^2 \cos^2 I / \sin^2 I}{
            2\epsilon \cos^2 I}
        = \frac{\epsilon}{\sin^2 I}.
\end{equation}
